Supplemental Materials for: Meta-analytic approaches and new effect sizes to account for heterogeneous effects in comparative physiology

### 2021-07-16 MS/supp.Rmd
Fixed ARR/ARRD code, removed VCV matrix because not working

---
title: "Supplemental Materials for: Meta-analytic approaches and new effect sizes to account for heterogeneous effects in comparative physiology"
author: Daniel W.A. Noble, Patrice Potter, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, tidy = TRUE)
options(digits=2)
```

@@

Daniel W.A. Noble, Patrice Potter, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa (2022) *Meta-analytic approaches and "new" effect sizes to account for heterogeneous effects in comparative physiology*. **Journal of Experimental Biology**

@@

```{r, class.source='klippy', results = 'hide'}
data_CS1a<-data_CS1a %>% 
              mutate(ARR=(mean_high-mean_low)/(acc_temp_high - acc_temp_low),
                     Var_ARR=((1/(acc_temp_high-acc_temp_low))^2*(sd_low^2/n_low_adj+sd_high^2/n_high_adj))) # Need to double check this formula, Shinichi pointed out that there might be a mistake.
```

In some cases, CTmax was measured twice on the same cohort of animals (e.g. using a different endpoint), which generates a source of non-independence between effect sizes. We can account for this source of non-independence by correlating errors between dependent effect sizes. The `make_VCV_matrix` from [`metaAidR`](https://github.com/daniel1noble/metaAidR) is a handy function to perform this correction, and assume a correlation of r = 0.5 by default. We termed the created variance-covariance matrix `VCV_ARR`.

```{r, class.source='klippy', results = 'hide', message=FALSE}
# Variance covariance matrix to account for non-independence of measurements performed on the same cohorts of animals
VCV_ARR<- metaAidR::make_VCV_matrix(data_CS1a, V = "Var_ARR", cluster = "shared_measurement_ID", obs = "es_ID")

# This method should be similar from clubSandwich
VCV_ARR2 <- clubSandwich::impute_covariance_matrix(data_CS1a$Var_ARR, cluster = "shared_measurement_ID", r =0.5)
```

@@

```{r, class.source='klippy'}
# Multi-level meta-analytic model
Model_CS1a_MLMA <- metafor::rma.mv(ARR ~ 1, V = VCV_ARR2, 
                   method="REML",
                   random=list(~1|species_ID,
                               ~1|es_ID), 
                   test="t",
                   dfs = "contain",
                   data=data_CS1a)
summary(Model_CS1a_MLMA)

# Calculate Prediction Intervals
PIs_CS1a_MLMA <- predict(Model_CS1a_MLMA)
print(PIs_CS1a_MLMA)

# Calculate I2
I2_CS1a <- orchaRd::i2_ml(Model_CS1a_MLMA)
print(I2_CS1a)
```

@@

@@

  mutate(ARR_f=((mean_high_f-mean_low_f)/(acc_temp_high-acc_temp_low))) %>% 
  mutate(ARR_m=((mean_high_m-mean_low_m)/(acc_temp_high-acc_temp_low))) %>% 
  mutate(ARRD=ARR_f - ARR_m) %>% 
  mutate(Var_ARRD=((1/(acc_temp_high-acc_temp_low))^2*(sd_low_f^2/n_low_f_adj+sd_high_f^2/n_high_f_adj+sd_low_m^2/n_low_m_adj+sd_high_m^2/n_high_m_adj)))  
```

@@

Again, we first started by accounting for the non-independence of CTmax estimates measured on the same cohorts using a variance-covariance matrix termed as `VCV_ARRD`.

```{r, class.source='klippy',  results = 'hide', message=FALSE}
# Variance covariance matrix to account for non-independence of measurements performed on the same cohorts of animals
#VCV_ARRD<- metaAidR::make_VCV_matrix(data_CS1b, V = "Var_ARR", cluster = "shared_measurement_ID", obs = "es_ID")
VCV_ARRD <- clubSandwich::impute_covariance_matrix(data_CS1b$Var_ARRD, cluster = "shared_measurement_ID", r =0.5)
```

@@

```{r, class.source='klippy'}
# Multi-level meta-analytic model
model_CS1b_MLMA<-metafor::rma.mv(yi= ARRD~ 1, V = VCV_ARRD, 
                   method="REML",
                   random=list(~1|species_ID,
                               ~1|es_ID), 
                   test="t",
                   dfs = "contain",
                   data=data_CS1b)
print(model_CS1b_MLMA)

# Calculate Prediction Intervals
PIs_CS1b_MLMA <- predict(model_CS1b_MLMA)
print(PIs_CS1b_MLMA)

# Calculate I2
I2_CS1b <- orchaRd::i2_ml(model_CS1b_MLMA)
print(I2_CS1b)
```

>>>>

---
title: "Supplemental Materials for: Meta-analytic approaches and new effect sizes to account for heterogeneous effects in comparative physiology"
author: Daniel W.A. Noble, Patrice Pottier, Samantha Burke, Szymon M. Drobniak, Malgorzata Lagisz, Rose E. O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, tidy = TRUE)
options(digits=2)
```

@@

Daniel W.A. Noble, Patrice Pottier, Sammantha Burke, Szymon M. Drobniak, Malgorzata Lagisz, Rose E. O’Dea, Shinichi Nakagawa (2022) *Meta-analytic approaches and "new" effect sizes to account for heterogeneous effects in comparative physiology*. **Journal of Experimental Biology**

@@

In this example, we will only use data measured on independent groups of animals, thus excluding data measured twice on the same cohort of animals (i.e. animals sharing the same `shared_measurement_ID` in the data). Note that it is possible to model this dependency and examples are provided in Noble et al. (2017) and Pottier et al. *in review*

# Calculate the effect sizes
data_CS1a<-data_CS1a %>% 
              mutate(ARR=(mean_high-mean_low)/(acc_temp_high - acc_temp_low),
                     Var_ARR=((1/(acc_temp_high-acc_temp_low))^2*(sd_low^2/n_low_adj+sd_high^2/n_high_adj)))

# Remove dependent effect sizes
data_CS1a<-filter(data_CS1a,
                    es_ID!="es85"&
                    es_ID!="es86"&
                    es_ID!="es87"&
                    es_ID!="es98"&
                    es_ID!="es282"&
                    es_ID!="es283"&
                    es_ID!="es284"&
                    es_ID!="es286") 
```

@@

```{r, class.source='klippy'}

# Multi-level meta-analytic model
model_CS1a_MLMA <- metafor::rma.mv(ARR ~ 1, V = Var_ARR, 
                   method="REML",
                   random=list(~1|species_ID,
                               ~1|es_ID), 
                   test="t",
                   data=data_CS1a)
print(model_CS1a_MLMA)

# Calculate Prediction Intervals
PI_CS1a_MLMA <- predict(model_CS1a_MLMA)
print(PI_CS1a_MLMA)

# Calculate I2
I2_CS1a <- orchaRd::i2_ml(model_CS1a_MLMA)
print(I2_CS1a)

# Plot the results
orchard_plot(model_CS1a_MLMA, mod="Int", xlab="ARR")
```

We can see from above that the model estimates an overall meta-analytic mean as `r coef(model_CS1a_MLMA)` with a 95\% CI from `r model_CS1a_MLMA$ci.lb` to `r model_CS1a_MLMA$ci.ub`. In other words, for each degree variation in acclimation temperature, the heat tolerance of the animals included in this data set vary by `r coef(model_CS1a_MLMA)` degrees on average. 95\% of the time we expect the mean in repeated samples to fall between `r PI_CS1a_MLMA$cr.lb` to `r PI_CS1a_MLMA$cr.ub`.
Our prediction intervals suggest that heterogeneity is high, reflected also in our $I_{total}^2$ estimate of `r I2_CS1a[1]*100`\%.  Approximately, `r I2_CS1a[2]*100`\% of the total heterogeneity is the result of differences between species and `r I2_CS1a[3]*100`\% of the total heterogeneity is associated with the residuals. 
With this effect size (slope), the 'nuisance heterogeneity' is accounted for, to some extent, at the level of the effect size.

@@

As before, we will only use data measured on independent groups of animals, thus excluding data measured twice on the same cohort of animals (i.e. animals sharing the same `shared_measurement_ID` in the data).

@@

  mutate(ARR_f=((mean_high_f-mean_low_f)/(acc_temp_high-acc_temp_low))) %>% 
  mutate(ARR_m=((mean_high_m-mean_low_m)/(acc_temp_high-acc_temp_low))) %>% 
  mutate(ARRD=ARR_f - ARR_m) %>% 
  mutate(Var_ARRD=((1/(acc_temp_high-acc_temp_low))^2*(sd_low_f^2/n_low_f_adj+sd_high_f^2/n_high_f_adj+sd_low_m^2/n_low_m_adj+sd_high_m^2/n_high_m_adj)))

# Remove dependent effect sizes
data_CS1b<-filter(data_CS1b,
                    es_ID!="es85"&
                    es_ID!="es86"&
                    es_ID!="es87"&
                    es_ID!="es98")
```

@@

@@

```{r, class.source='klippy'}

# Multi-level meta-analytic model
model_CS1b_MLMA<-metafor::rma.mv(yi= ARRD~ 1, V = Var_ARRD, 
                   method="REML",
                   random=list(~1|species_ID,
                               ~1|es_ID), 
                   test="t",
                   data=data_CS1b)
print(model_CS1b_MLMA)

# Calculate Prediction Intervals
PI_CS1b_MLMA <- predict(model_CS1b_MLMA)
print(PI_CS1b_MLMA)

# Calculate I2
I2_CS1b <- orchaRd::i2_ml(model_CS1b_MLMA)
print(I2_CS1b)

# Plot the results
orchard_plot(model_CS1b_MLMA, mod="Int", xlab="ARR")
```

We can see from above that the model estimates an overall meta-analytic mean as 0.002 with a 95\% CI from `r model_CS1b_MLMA$ci.lb` to `r model_CS1b_MLMA$ci.ub`. In other words, for each degree variation in acclimation temperature, the heat tolerance of females vary by an average of 0.002 degrees more than males. Note that the confidence intervals overlap with zero, which indicates that the difference between females and males is small, and not statistically significant. 95\% of the time we expect the mean in repeated samples to fall between `r PI_CS1b_MLMA$cr.lb` to `r PI_CS1b_MLMA$cr.ub`.
Our prediction intervals suggest that heterogeneity is small to moderate, reflected also in our $I_{total}^2$ estimate of `r I2_CS1b[1]*100`\%. Approximately, `r I2_CS1b[2]*100`\% of the total heterogeneity is the result of differences between species and `r I2_CS1b[3]*100`\% of the total heterogeneity is associated with the residuals.
